How environmental drivers of synchrony interact

### 2022-01-14 Paper/Paper.Rnw
working on Methods and Results

%***Tentative author order, Bell last for his data contributions
\author[a,b,*]{Daniel C. Reuman}
\author[c]{Max Castorani}
\author[d]{Kyle Cavanaugh}
\author[e]{Lawrence W. Sheppard}
\author[c]{Jonathan Walter}
\author[f,g]{Tom Bell}

\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}
\affil[c]{Department of Environmental Sciences, University of Virginia}
\affil[d]{Department of Geography, University of California, Los Angeles}
\affil[e]{Marine Biological Association of the United Kingdom}
\affil[f]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
\affil[g]{Earth Research Institute, University of California, Santa Barbara}
\affil[*]{Corresponding author, reuman@ku.edu}

@@

\section{General theory}

@@

%Description of data after cleaning
The core of this paper is the new theory, but to help illustrate that theory, we also apply
it to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
associated environmental measurements. 
The data are based on a subset of those used by \cite{CastoraniSheppardKelp} and
\cite{Walter2021}, and data details are given in those papers and in SI section \ref{SM-sect:data}. We here
summarize the format of the data after a preparation and cleaning process was implemented. 
\cite{CastoraniSheppardKelp} and \cite{Walter2021} also provide an introduction to 
kelp ecology, and information on why giant kelp is an excellent species 
for studies of synchrony. After preparation, kelp data consisted of
$\Sexpr{length(SBlocstouse)+length(CC1locstouse)+length(CC2locstouse)}$ quarterly kelp abundance time series
from locations along the CA coast, each time series spanning from quarter $\Sexpr{quarters[1,3]}$ 
of $\Sexpr{quarters[1,2]}$ to quarter $\Sexpr{quarters[dim(quarters)[1],3]}$ of 
$\Sexpr{quarters[dim(quarters)[1],2]}$.   
Time series were grouped into three regions which were analyzed separately: a more northerly
central CA group of $\Sexpr{length(CC1locstouse)}$ locations (called CCA1); a more southerly central CA group of 
$\Sexpr{length(CC2locstouse)}$ locations (CCA2); and a group of $\Sexpr{length(SBlocstouse)}$ locations 
from southern CA, close to Santa Barbara
(SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of mean quarterly kelp biomass per unit 
useable habitat in a 500m stretch of coastline. We used coastline segments where kelp was persistent
through essentially all of the $\Sexpr{quarters[1,2]}$-$\Sexpr{quarters[dim(quarters)[1],2]}$ 
period (SI section \ref{SM-sect:data}). We also had estimates
of maximum wave height and mean nitrate concentration for each quarter and location; both waves
and nutrients are known to influence kelp population dynamics and synchrony 
\citep{Cavanaugh2013,CastoraniSheppardKelp}.

%Tentative broad outline
%2) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated,
%about the lags we used, and justification for them etc. Might also need to cite a sup mat section. Maybe 
%we need one single sup mat section called "supplementary methods" or "details of methods" or something. Actually
%let's have two of these, one for data, one for stats
%3) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
%other parts of the main theoretical equation were estimated. Basically it's just about how to do spectral
%estimates of spectra and cross spectra, and then rest follows from the main theoretical equation.
%4) Somehow need to get across what was done - different parts of main equation were examined to explore the different
%aspects of how the mechanism works for the focal taxon. Elaborate as appropriate.

Data for the project are archived at X.
%***DAN: Need to get the data location from Jon, I think he posted it, right? Maybe he did not put it all i
%one place. Check with him. At the end of the Max/Lawrence paper, several data locations are listed.
%I have an email out to Jon and Max.
All computations were done in R version 3.4.4 on an ordinary laptop which was running Ubuntu Linux
16.04. Complete codes and data are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.

@@

%Para basically confirming earlier work, as a starter
Our theoretical results demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
can depend strongly on timescale, even having different signs on different timescales. 
First, for all of our theoretical case studies, the magnitude of 
interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
compare the dashed and solid lines). So interactions can contribute substantially to overall population 
synchrony. Second, in contrast to the direct Moran effects of a 
single environmental variable, which provide non-negative 
contributions to population synchrony whenever, as is typical, environmental synchrony 
is positive, interactions between Moran effects can be positive or negative. For instance, case study A
showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and 
case study C showed negative interactions at
long timescales (Fig. \ref{fig:TheoryCaseA}i,m), or, for some parameters, at all timescales 
(Fig. \ref{fig:TheoryCaseA}m).
Thus interaction effects can either augment or reduce population synchrony from what its value would be 
if only direct Moran effects operated. Finally, interactions between Moran effects depended strongly on timescale 
for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and differed in sign across timescales for case studies
A and B and for one of the C scenarios. This augments earlier studies demonstrating that direct Moran effects can 
depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The points made in this paragraph
also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
straightforwardly, and those papers are case studies, so our results provide novel theoretical generality.
Our theory can also be interpreted to provide intuition about how interacting Moran effects work,
as follows.

@@

Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Interactions between Moran effects were 
negative on the shortest timescales 
(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel) because the 
population influences of Moran effects of timescale 2
quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
2 quarters, producing negative interactions. 
Interaction effects on long timescales were instead synergistic, i.e., positive, for B 
(Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
At timescale 3 quarters, Moran effects are reinforcing because the lag equals the timescale,
producing Moran effects of the two drivers which are in phase 
(Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
makes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
interactions. Thus interactions between Moran drivers were zero 
at timescales 1 and 3 years for B, were 
negative for short timescales and for timescale 1.5 years, and were positive for 
timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
Theory captures these complexities because 
the size of the 3-quarter phase lag of this scenario, relative to a timescale,
is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B (SI section \ref{SM-sect:details_case_studies}),
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as that of
$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}g),
which controls the sign of the interaction term,
$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, in our 
theory (\ref{rhoww_T1to3}).

Cases A and B assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
(SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
but case study C illustrates what can happen when noise processes have periodic components, e.g., 
both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
between the lags of their effects on populations, as we now explain. 
For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
$\epsilon^{(1)}$ effects on populations were positive, and were 
lagged (i.e., delayed in their influence) by 1 quarter; but 
$\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
For scenario 1 of C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
1 quarter. Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
were exactly 2 quarters offset from the annual negative effects of $\epsilon^{(2)}$, 
producing reinforcing interactions 
between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
1 quarter. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
coincided with the negative effects of $\epsilon^{(2)}$ every year, a half-phase shift for an 
annually oscillating effect, which corresponds to negative interactions between Moran influences
and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 
is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
Our theory captures these complexities
because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
$\epsilon^{(2)}$ compared to a given timescale, while 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
therefore captures whether relative lags of effects and lags between 
the variables themselves align or misalign, with the degree of alignment
reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
which is half the interaction term in our theory (\ref{rhoww_T1to3}).
On long timescales, interactions were the
same in the two scenarios because, on those timescales, sub-annual lags 
make negligible difference. Interactions were then negative 
because the effects of $\epsilon^{(1)}$ on populations were 
positive and those of $\epsilon^{(2)}$ were negative.

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Theory/IntroTheory.pdf}
\includegraphics[width=.5\textwidth]{../Results/Theory/PedagogFig.pdf}
\caption{Example and basic concept of interacting Moran effects. The example (a), which 
is based on a simple model described in the Introduction, shows that 
interaction effects are possible, and can be substantially positive or negative. 
We used $\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})=\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})=0.7$,
and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
other and to a parameter, $\phi$. For (b) and (c), which represent distinct scenarios, 
the top sinusoid on each panel represents 
the component of oscillation at a given timescale of an environmental variable, $\epsilon_i^{(1)}$,
measured in a sampling location $i$, and the bottom sinusoid represents the 
component at the same timescale for another environmental variable, $\epsilon_i^{(2)}$; the 
oscillations are lagged from each other by an amount $l_n$. Environmental effects on 
the population, $w_i(t)$, in location $i$ are also lagged, by amounts $l_{e1}$ and
$l_{e2}$. For this setup, a positive interaction between Moran effects can occur on the
given timescale when lags align, i.e, $l_{e1}=l_{e2}+l_n$, as in panel (a), so that environmental
effects on the population coincide and reinforce each other. All environmental effects on 
populations are assumed positive in these examples. A negative
effect can occur when lags produce out-of-phase influence of environmental variables on
populations, i.e., $l_{e1}$ differs from $l_{e2}+l_n$ by half the given timescale/period,
as in panel (b), so that the distinct influences tend to work against each other in 
bringing about population oscillations on the given timescale.  
A timescale-specific approach is essential to this reasoning.}\label{fig:IntroTheory}
\end{figure}

@@

\begin{figure}
\caption{To fill.}\label{fig:map}
\end{figure}

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\caption{Filler.}\label{fig:indirectMoran}
\end{figure}

>>>>

%***Tentative author order, Bell last for his data contributions
\author[a,*]{Daniel C. Reuman}
\author[b]{Max Castorani}
\author[c]{Kyle Cavanaugh}
\author[d]{Lawrence W. Sheppard}
\author[b]{Jonathan Walter}
\author[e,f]{Tom Bell}

\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
\affil[b]{Department of Environmental Sciences, University of Virginia}
\affil[c]{Department of Geography, University of California, Los Angeles}
\affil[d]{Marine Biological Association of the United Kingdom}
\affil[e]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
\affil[f]{Earth Research Institute, University of California, Santa Barbara}
\affil[*]{Corresponding author, reuman@ku.edu}

@@

\section{General Theory}

@@

%Description of data after cleaning
The core of this paper is the new theory, but to help illustrate that theory, fulfilling goal G3 of the 
Introduction, we also apply
the theory to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
associated environmental measurements. 
The data are based on a subset of those used by \cite{CastoraniSheppardKelp} and
\cite{Walter2021}, and data details are given in those papers and in SI section \ref{SM-sect:data}. We here
summarize the format of the data after a preparation and cleaning process was implemented. 
\cite{CastoraniSheppardKelp} and \cite{Walter2021} also provide an introduction to 
kelp ecology, and information on why giant kelp is an excellent species 
for studies of synchrony. After preparation, kelp data consisted of
$\Sexpr{length(SBlocstouse)+length(CC1locstouse)+length(CC2locstouse)}$ quarterly kelp abundance time series
from locations along the CA coast, each time series spanning from quarter $\Sexpr{quarters[1,3]}$ 
of $\Sexpr{quarters[1,2]}$ to quarter $\Sexpr{quarters[dim(quarters)[1],3]}$ of 
$\Sexpr{quarters[dim(quarters)[1],2]}$.   
Time series were grouped into three regions which were analyzed separately: a more northerly
central CA group of $\Sexpr{length(CC1locstouse)}$ locations (called CCA1); a more southerly central CA group of 
$\Sexpr{length(CC2locstouse)}$ locations (CCA2); and a group of $\Sexpr{length(SBlocstouse)}$ locations 
from southern CA, close to Santa Barbara
(SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of mean quarterly kelp biomass per unit 
useable habitat in a 500m stretch of coastline. We used coastline segments where kelp was persistent
through essentially all of the $\Sexpr{quarters[1,2]}$-$\Sexpr{quarters[dim(quarters)[1],2]}$ 
period (SI section \ref{SM-sect:data}). We also had estimates
of maximum wave height and mean nitrate concentration for each quarter and location; both waves
and nutrients are known to influence kelp population dynamics and synchrony 
\citep{Cavanaugh2013,CastoraniSheppardKelp}.

Coefficients of the model (\ref{main_model_1})-(\ref{main_model_4}) were separately selected for each of our three regions 
using linear regression methods and building on extensive prior work on the drivers of kelp dynamics. A no-intercept 
regression model of kelp at time $t$ in location $i$, i.e., 
$w_i(t)$ in (\ref{main_model_1}), against lagged values of kelp 
($w_i(t-l)$ for $l=1,\ldots,n$ in (\ref{main_model_1})), lagged and unlagged values of nitrates
($\epsilon_i^{(1)}(t-l)$ for $l=0,\ldots,m_1$ in (\ref{main_model_2})), and values of waves
($\epsilon_i^{(2)}(t-l)$ for $l=0,\ldots,m_2$ in (\ref{main_model_3})) was fitted using standard regression methods.
The same regression coefficients were used for all locations within a region.
Waves influence kelp dynamics through direct disturbance events which can damage kelp or extirpate kelp locally 
when waves are large \citep{Dayton1992,Cavanaugh2011,Bell2015}. Thus wave effects 
are immediate and $m_2=0$ was used. Nitrates are known to fuel rapid kelp growth, though in some areas
effects appear delayed by about one quarter because our kelp data quantify canopy (surface) biomass, and it can
take time for damaged kelp to grow back to the surface 
\citep{Zimmerman1986,Cavanaugh2011,Bell2015}. 
Therefore $m_1=1$ was used. Kelp holdfasts on the sea floor can last for multiple years, so kelp lag effects were considered:
we used $n=4, 8, 12$ in separate analysis, with this choice making no substantive difference to results (see Results).
Regression coefficient determined the quantities $b_1,\ldots,b_n$, $p_0^{(1)}$, $p_1^{(1)}$, and $p_0^{(2)}$
in (\ref{main_model_1})-(\ref{main_model_3}), and therefore the quantities $f_{\mcB}$, $f_{\mcP^{(1)}}$ and
$f_{\mcP^{(2)}}$ in (\ref{rhoww_T1to3}), for each of our three regions (see General Theory).

To estimate the components $\rho_{ww}$, $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and 
$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was detrended kelp time series in one of our regions, 
$\epsilon^{(1)}$ was detrended nitrate time
series in the region, and $\epsilon^{(2)}$ was detrended wave disturbance time series in the region, 
we applied the definitions of these quantities (see General Theory), which required estimation from data of
the spectra and cross spectra $S_{w_i w_j}$ and $S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}$ for $i,j=1,\ldots,N$
and $a,b=1,2$, where $N$ is the number of locations in the region being considered. Spectral quantities were
computed using the consistent estimator of section 7.4 of \cite{Brillinger2001}. 
%***DAN: check this section number, and if possible give a specific formula of theorem or something, and also 
%check the below text and maybe augment it
The estimator is a smoothed periodogram, with the width of the smoothing kernel selected to 
increase with the square root of time series length. Theory was interpreted in relation to kelp ecology and
theoretical case studies by plotting the components of (\ref{rhoww_T1to3}) for each of our regions.

Data for the project are publicly archived \citep{SBCLTER2021}.
All computations were done in R version 3.4.4 on an ordinary laptop which was running Ubuntu Linux
16.04. Complete codes for the entire project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.

@@

%Para basically confirming earlier work, as a starter
Helping fulfill goal G2 of the Introduction, our theoretical results 
demonstrate that interaction effects between Moran drivers: 1) can be 
comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
can depend strongly on timescale, even having different signs on different timescales. 
First, for all of our theoretical case studies, the magnitude of 
interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
compare the dashed and solid lines). So interactions can contribute substantially to overall population 
synchrony. Second, in contrast to the direct Moran effects of a 
single environmental variable, which provide non-negative 
contributions to population synchrony whenever, as is typical, environmental synchrony 
is positive, interactions between Moran effects can be positive or negative. For instance, case study A
showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and 
case study C showed negative interactions at
long timescales (Fig. \ref{fig:TheoryCaseA}i,m), or, for some parameters, at all timescales 
(Fig. \ref{fig:TheoryCaseA}m).
Thus interaction effects can either augment or reduce population synchrony from what its value would be 
if only direct Moran effects operated. Finally, interactions between Moran effects depended strongly on timescale 
for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and differed in sign across timescales for case studies
A and B and for one of the C scenarios. This augments earlier studies demonstrating that direct Moran effects can 
depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The points made in this paragraph
also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
straightforwardly, and those papers are case studies, so our results provide novel theoretical generality.
Our theory can also be interpreted to provide intuition about how interacting Moran effects work,
as follows.

@@

%Case syiudy B
Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Interactions between Moran effects were 
negative on the shortest timescales 
(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel) because the 
population influences of Moran effects of timescale 2
quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
2 quarters, producing negative interactions. 
Interaction effects on long timescales were instead synergistic, i.e., positive, for B 
(Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
At timescale 3 quarters, Moran effects are reinforcing because the lag equals the timescale,
producing Moran effects of the two drivers which are in phase 
(Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
makes the Moran effects of 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
interactions. Thus interactions between Moran drivers were zero 
at timescales 1 and 3 years for B, were 
negative for short timescales and for timescale 1.5 years, and were positive for 
timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
Theory captures these complexities because 
the size of the 3-quarter phase lag of this scenario, relative to a timescale,
is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B (SI section \ref{SM-sect:details_case_studies}),
the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as that of
$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}g),
which controls the sign of the interaction term,
$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, in our 
theory (\ref{rhoww_T1to3}).

%Case study C
Cases A and B assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
(SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
but case study C illustrates what can happen when noise processes have periodic components, e.g., 
both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
between the lags of their effects on populations, as we now explain. 
For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
$\epsilon^{(1)}$ effects on populations were positive, and were 
lagged (i.e., delayed in their influence) by 1 quarter; but 
$\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
For scenario 1 of C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
1 quarter. Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
were exactly 2 quarters offset from the annual negative effects of $\epsilon^{(2)}$, 
producing reinforcing interactions 
between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
1 quarter. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
coincided with the negative effects of $\epsilon^{(2)}$ every year, a half-phase shift for an 
annually oscillating effect, which corresponds to negative interactions between Moran influences
and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 
is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
Our theory captures these complexities
because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
$\epsilon^{(2)}$ compared to a given timescale, while 
$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
therefore captures whether relative lags of effects and lags between 
the variables themselves align or misalign, with the degree of alignment
reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
which is half the interaction term in our theory (\ref{rhoww_T1to3}).
On long timescales, interactions were the
same in the two scenarios because, on those timescales, sub-annual lags 
make negligible difference. Interactions were then negative 
because the effects of $\epsilon^{(1)}$ on populations were 
positive and those of $\epsilon^{(2)}$ were negative.

%Paralels the first para, but for kelp
We now apply our theory to kelp, fulfilling goal G3 of the Introduction. Kelp results confirm
the earlier theoretical results, based on our case studies 
(see above text and Fig. \ref{fig:TheoryCaseA}a,e,i,m),
that interaction effects between Moran drivers: 1) can be comparble in strength to direct Moran
effects; 2) can be either synergistic or destructive; and 3) can 
depend strongly on timescale, even having different signs on different timescales.
First, for all of our regions and for essentially all timescales, the magintude of interaction 
effects was comparable to that of direct Moran effects (Fig. \ref{fig:SummaryKelpFig}
for the CCA1 and SB regions and Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the CCA2 region).
Second, interaction effects could be positive (e.g., annual timescales, CCA1 region on Fig. 
\ref{fig:SummaryKelpFig}a), or negative (e.g., annual timescales, SB region on Fig. 
\ref{fig:SummaryKelpFig}d, or long timescales $>8$y for either region, Fig. 
\ref{fig:SummaryKelpFig}c,f). Finally, interactions depended on timescale, e.g., for the CCA1 
region they were positive on annual timescales (Fig. \ref{fig:SummaryKelpFig}a) and
negative on timescales $>8$y (Fig. \ref{fig:SummaryKelpFig}c).

The way direct Moran effects work in our theory is fairly straightforward. 
For instance, the direct Moran effect
term in (\ref{rhoww_T1to3}) for nitrates, i.e., 
$|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}/|f_{\mcB}|^2$, 
equals the synchrony of the nitrate time series themselves, i.e.,
$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$, times the 
timescale-specific strength of influence of nitrates on kelp, 
i.e., $|f_{\mcP^{(1)}}|^2$, and modulated
by a term quantifying the autoregressive nature of kelp dynamics,
$1/|f_{\mcB}|^2$. These components are plotted in Fig. \ref{fig:directMoran} (CCA1 and SB regions) and 
Fig. \ref{SM-fig:directMoran_SM_lag4} (CCA2 region) so their 
effects on synchrony can be assessed across timescales and regions. 
Both nitrates and waves fluctuate seasonally in CA. Thus nitrate and wave 
synchrony had a strong annual component (Fig. \ref{fig:directMoran}b,d,q,s), which produced 
some of the annual synchrony observed in kelp 
(Fig. \ref{fig:SummaryKelpFig}a,d, green and blue lines).
Nitrates and waves are also synchronous on long timescales $>8$y in CA 
(Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation [NPGO; \cite{CastoraniSheppardKelp}]. The long-timescale synchrony
in nitrates and waves produced some of the long-timescale
synchrony observed in kelp (Fig. \ref{fig:SummaryKelpFig}c,f, green and blue lines).
Kelp synchrony was generally stronger in the CCA1 region compared to the SB region 
(compare Fig. \ref{fig:SummaryKelpFig}a-c to d-f) in part because the synchrony of nitrates, and of 
waves, was more pronounced in the CCA1 region than in the SB region 
(compare Fig. \ref{fig:directMoran}b,g,l to q,v,aa, and 
Fig. \ref{fig:directMoran}d,i,n to s,x,cc), and also because waves
had a stronger influence on kelp in CCA1 compared to SB (compare 
Fig. \ref{fig:directMoran}c,h,m to r,w,bb).
Regression coefficients which determined $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
and $f_{\mcB}$ (see Methods) are in Table \ref{SM-tab:regcoefs}.

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Theory/IntroTheory.pdf}
\includegraphics[width=.5\textwidth]{../Results/Theory/PedagogFig.pdf}
\caption{Example and basic concept of interacting Moran effects. The example (a), which 
is based on a simple model described in the Introduction, shows that 
interaction effects are possible, and can be substantially positive or negative. 
We used $\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})=\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})=0.7$,
and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
other and to a parameter, $\phi$. For (b) and (c), which represent distinct scenarios, 
the top sinusoid on each panel represents 
the component of oscillation at a given timescale of an environmental variable, $\epsilon_i^{(1)}$,
measured in a sampling location $i$, and the bottom sinusoid represents the 
component at the same timescale for another environmental variable, $\epsilon_i^{(2)}$; the 
oscillations are lagged from each other by an amount $l_n$. Environmental effects on 
the population, $w_i(t)$, in location $i$ are also lagged, by amounts $l_{e1}$ and
$l_{e2}$. For this setup, a positive interaction between Moran effects can occur on the
given timescale when lags align, i.e, $l_{e1}=l_{e2}+l_n$, as in panel (a), so that environmental
effects on the population coincide and reinforce each other. All environmental effects on 
populations are assumed positive in these examples. A negative
effect can occur when lags produce out-of-phase influence of environmental variables on
populations, i.e., $l_{e1}$ differs from $l_{e2}+l_n$ by half the given timescale/period,
as in panel (b), so that the distinct influences tend to work against each other in 
bringing about population oscillations on the given timescale.  
A timescale-specific approach is essential to this reasoning. See the online version for a color
rendering of all figures.}\label{fig:IntroTheory}
\end{figure}

@@

\begin{figure}
\caption{To fill. See the online version for a color
rendering of all figures.}\label{fig:map}
\end{figure}

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\caption{Filler. See the online version for a color
rendering of all figures.}\label{fig:indirectMoran}
\end{figure}
